---
title: "TG_Assignment2"
author: "Tamar Gazit (B00842040)"
format: html
editor: visual
---

## Data Cleaning

```{r}
# Load necessary libraries
library(tidyverse) # For data wrangling
library(dplyr) # For data manipulation
library(haven) # For reading SPSS (.sav) files
library(flexplot) # For GLM analysis
library(patchwork) # To arrange multiple flexplots into a single layout
library(apaTables) # For table generation
library(car) # For multicollinearity check

# Import the dataset
data <- read_sav("~/Desktop/PSYR 6003 - Advanced Statistics/TG_Assignment2/PSYR6003.A2.sav")
```

Change 'sex' to a factor variable with 0 = "Female" and 1 = "Male":

```{r}
# Check unique values in the sex column
unique(data$sex)

# Remove cases where sex is "Other (specify)"
data <- data %>%
  filter(sex %in% c("Female", "Male"))

# Check if the cases were successfully removed
unique(data$sex)  # Should now only show "Female" and "Male"

# Convert sex to a factor with 0 = Female, 1 = Male
data$sex <- factor(data$sex, 
                   levels = c("Female", "Male"),  # Keep order consistent
                   labels = c(0, 1))  # Assign 0 to Female, 1 to Male

# Verify changes
table(data$sex)  # Check how many participants are in each group
```

Create three new variables (subscale scores):

```{r}
# 1. REVERSE-CODE the original (unmodified) tipm.CONS2.3y item (on 1–7 scale)
data <- data %>%
  mutate(tipm.CONS2.3y_rc = 8 - tipm.CONS2.3y)

# 2. COMPUTE SUBSCALE MEANS
data <- data %>%
  mutate(
    # Socially Prescribed Perfectionism: Mean of 5 items
    SPP_total = rowMeans(select(., mps.SPP1.3y, mps.SPP2.3y, mps.SPP3.3y, mps.SPP4.3y, mps.SPP5.3y), na.rm = TRUE),
    
    # Conscientiousness: Use reverse-coded version
    CONS_total = rowMeans(select(., tipm.CONS1.3y, tipm.CONS2.3y_rc), na.rm = TRUE),
    
    # Negative Affect: Mean of 12 items
    NA_total = rowMeans(select(., 
      guilt1.3y, guilt2.3y, guilt3.3y,
      dep1.3y, dep2.3y, dep3.3y,
      fear1.3y, fear2.3y, fear3.3y,
      host1.3y, host2.3y, host3.3y
    ), na.rm = TRUE)
  )

# Summary of total scores
summary(select(data, SPP_total, CONS_total, NA_total))

# Check for missing values
colSums(is.na(select(data, SPP_total, CONS_total, NA_total)))
```

## 1. State Hypotheses

This study tested two hypotheses:

First (H1), it was predicted that sex, conscientiousness, and socially prescribed perfectionism (SPP) would each significantly predict negative affect. Specifically, women were expected to report higher negative affect than men, higher conscientiousness was expected to relate to lower negative affect, and higher SPP was expected to relate to higher negative affect.

Second (H2), it was predicted that SPP would explain unique variance in negative affect beyond the effects of sex and conscientiousness.

## 2. Visualize Data

#### A. Univariate Distributions

```{r}
# Visualize outcome variable: Negative Affect
p1 <- flexplot(NA_total ~ 1, data = data)

# Visualize predictors: Conscientiousness, SPP, and Sex
p2 <- flexplot(CONS_total ~ 1, data = data)
p3 <- flexplot(SPP_total ~ 1, data = data)
p4 <- flexplot(sex ~ 1, data = data)

# Display all 4 in a patchwork grid
(p1 | p2) / (p3 | p4)
```

#### B. Bivariate Relationships

```{r}
# Negative Affect by Conscientiousness
b1 <- flexplot(NA_total ~ CONS_total, data = data)

# Negative Affect by SPP
b2 <- flexplot(NA_total ~ SPP_total, data = data)

# Negative Affect by Sex
b3 <- flexplot(NA_total ~ sex, data = data)

# Display all 3 together
b1 + b2 + b3
```

## 3. Assumption Checking

```{r}
# Fit the full model (which predics Negative Affect from Sex, Conscientiousness, and SPP)
full_model <- lm(NA_total ~ sex + CONS_total + SPP_total, data = data)

# Visual diagnostic plots
visualize(full_model, plot = "residuals")

# Check multicollinearity
vif(full_model)
```

## 4. Test H1

```{r}
# Full model with all predictors
full_model <- lm(NA_total ~ sex + CONS_total + SPP_total, data = data)

# Manually extract estimates
summary(full_model)
estimates(full_model)
```

## 5. Test H2

```{r}
# Fit the reduced model (without SPP)
reduced_model <- lm(NA_total ~ sex + CONS_total, data = data)

# Compare models
model.comparison(reduced_model, full_model)

# Manually extract estimates
summary(reduced_model)
summary(full_model)
estimates(reduced_model)
estimates(full_model)
summary(full_model)$r.squared - summary(reduced_model)$r.squared
```

## 6. Create Tables

```{r}
# Table 1: Descriptives and Correlation Table
# =============================================
# Select analysis variables (excluding sex because it's categorical)
analysis_vars <- data %>%
  select(NA_total, CONS_total, SPP_total)

# Create the APA-formatted correlation table
apa.cor.table(analysis_vars, filename = "Table1_Descriptives_and_Correlations.doc")

# Table 2: Regression Coefficients and Model Summary
# =====================================================
# Run full model
full_model <- lm(NA_total ~ sex + CONS_total + SPP_total, data = data)

# Create regression table
apa.reg.table(full_model, filename = "Table2_Regression_Model.doc")
```
